---
title: "Targeted enrichment and discovery of jumbo phages"
author: "Jeffrey Blanchard"
date: "5/4/2023"
output:
  html_document:
    toc: true
    toc_depth: 4
    theme: united
    highlight: tango
---

# This report is an analysis of the Barre Woods filter metagenomes from the individual assembles.  Assembly, binning, quality control and classification was done on KBase https://narrative.kbase.us/narrative/145959  . The checkv and virsorter files were downloaded to this project folder.

## Load Libraries

```{r}
# Load libraries
library(tidyverse)
library(DT)
```

## Import Data

### Import vsorter files. Filter to include just the dsDNAphages

```{r}
virsorter <- read_tsv("data_individual_assembly/all-ind-virsorter-final-viral-score.tsv") %>% 
  filter(max_score_group == "dsDNAphage") %>% 
  select(-c(dsDNAphage, ssDNA, NCLDV, lavidaviridae)) %>% 
  rename("contig_id" = "seqname")
```

```{r}
checkv <- read_tsv("data_individual_assembly/all-ind-checkv_quality_summary.tsv") 
```


### Join checkv and virsorter dataframes

```{r}
virsorter_checkv <- left_join(virsorter, checkv, by = "contig_id")
```

### filter to high quality and complete

```{r}
virsorter_checkv_HQ <- virsorter_checkv %>% 
  filter(checkv_quality == "Complete" | checkv_quality == "High-quality")
```

```{r, eval=FALSE}
# write summary file
write_tsv(virsorter_checkv, "data_individual_assembly/all-ind-virsorter_checkv.tsv") 
write_tsv(virsorter_checkv_HQ, "data_individual_assembly/all-ind-virsorter_checkv_HQ.tsv") 
```

## Data analysis

### Barchart of checkv_quality
```{r}
virsorter_checkv %>% 
ggplot(aes(x = checkv_quality)) + 
  geom_bar() 
```

### Table with % checkv_quality in each category
```{r}
virsorter_checkv %>% 
  mutate(across(checkv_quality, as_factor)) %>% 
  group_by(checkv_quality) %>% 
  summarise(n = n()) %>%
  mutate(freq = n / sum(n))
```


### Histogram

```{r}
virsorter_checkv %>% 
ggplot(aes(x = length, fill = checkv_quality)) + 
  geom_histogram(colour = "black",  binwidth=10000) +
  ggtitle("Genome size of Phage") +
  xlab("Genome size") + 
  theme(text = element_text(size = 20, color="black"))
  theme(axis.text.x = element_text(angle = 90)) 
```

```{r}
virsorter_checkv_HQ %>% 
ggplot(aes(x = length, fill = checkv_quality)) + 
  geom_histogram(colour = "black",  binwidth=10000) +
  ggtitle("Genome size of Phage") +
  xlab("Genome size") + 
  #theme(text = element_text(size = 20, color="black"))
  theme(axis.text.x = element_text(angle = 45)) 
```
